---
title: "Time Series Analysis"
author: "David Faria"
date: "18 octobre 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Taken from (https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/)

## Importing data

```{r}
data("AirPassengers")

class(AirPassengers)
```

This is the start of the time series

```{r}
start(AirPassengers)
```

This is the end of the time series

```{r}
end(AirPassengers)
```

The cycle of this time series is 12months in a year

```{r}
frequency(AirPassengers)
```

The number of passengers are distributed across the spectrum

```{r}
summary(AirPassengers)
```


```{r}
plot(AirPassengers)
abline(reg=lm(AirPassengers~time(AirPassengers)))
```

This will print the cycle across years.

```{r}
cycle(AirPassengers)
```

This will aggregate the cycles and display a year on year trend

```{r}
plot(aggregate(AirPassengers,FUN=mean))
```

Box plot across months will give us a sense on seasonal effect

```{r}
boxplot(AirPassengers~cycle(AirPassengers))
```

## Important Inferences

* The year on year trend clearly shows that the #passengers have been increasing without fail.
* The variance and the mean value in July and August is much higher than rest of the months.
* Even though the mean value of each month is quite different their variance is small. Hence, we have strong seasonal effect with a cycle of 12 months or less.


# ARMA Time Series Modeling

* AR stands for auto-regression 
* MA stands for moving average

AR or MA are not applicable on non-stationary series.

## AR

$x(t) = alpha *  x(t - 1) + error (t)$

This equation is known as AR(1) formulation. The numeral one (1) denotes that the next instance is solely dependent on the previous instance.  The alpha is a coefficient which we seek so as to minimize the error function. Notice that x(t- 1) is indeed linked to x(t-2) in the same fashion. Hence, any shock to x(t) will gradually fade off in future.

## MA

$x(t) = beta *  error(t-1) + error (t)$

In MA model, noise / shock quickly vanishes with time. The AR model has a much lasting effect of the shock.